\begin{center}
{\bf \large A C K N O W L E D G E M E N T S}
\end{center}
\rule{\linewidth}{2mm} \pagestyle{empty} \vspace{0.25in}

~~~~The satisfaction that accompanies the successful completion of any work would be incomplete without the mention of the persons who made it possible, whose constant guidance and encouragement crown all the efforts. An understanding of the study like this, is never the outcome of the efforts of a single person, rather it bears the imprint of a number of persons who directly or indirectly helped me in the partial fulfillment of the project. I would be failing in my duty if I don't say a word of thanks to all these people.

First and foremost, I would like to thank Prof. Kandarpa Kumar Sarma, my supervisor without whose guidance, support and mentoring, I would not be able to complete this task. I must call myself lucky to be able to work under the supervision of such a supportive and friendly person.

Besides my supervisor, would like to thank Dr. Kumaresh Sarmah, assistant professor, Department of Electronics and Communication Technology, Gauhati University for his guidance and motivation throughout these years. I thank Dr. Pranjal Borah, assistant professor, Department of Instrumentation, Gauhati University for his guidance and suggestions from time to time.

I thank Prof. Tulshi Bezboruah, Dr. Navajit Saikia, Dr. Hidam Kumarjit Singh, Dr. Ram Kishore Roy, Dr. Hirendra Das for their motivation and support throughout the years of my Ph. D.

I extend my gratitude to all teaching and non-teaching members of the department of Electronics and Communication Engineering, Gauhati University including Mr. Manash Pratim Sarma, Dr. Subhash Chandra Rajbongshi, Dr. Jyoti Prakash Medhi, Dr. Aradhana Mishra, Mr. Anjan Talukdar, Mr. Apurba Vikash Kalita, Mr. Hirakjyoti Goswami, Mr. Amitava Dey for their valuable support during this period. I would specifically thank all the other research scholars of the department, including Mr. Manasjyoti Bhuyan, Mr. Amlan Jyoti Das, Dr. Mridusmita Sharma, Ms. Simantika Choudhury, Mr. Surajit Deka, Mr. Mrinmoy Sandilya for their motivation and support.

Finally, I would like to thank my parents and all my family members for their motivation and support.


\vspace*{11mm}





\goodbreak
\begin{multicols}{2}
\begin{flushleft}
Date: 01-01-2022 \\
Place: Gauhati University \\
\end{flushleft}

\begin{flushright}
...................................... \\
Sivaranjan Goswami \\
Enrolment No.- PA-181-839-0003\\
\end{flushright}
\end{multicols}
